biological data preperation

### 2018-09-19 src/biological_data_prep.Rmd
amo prev

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

@@

amo_yrmt.sort <- amo_yrmt[order(amo_yrmt$year),]
amo_yrmt
```
rename time to month.. and value to amo_sample
```{r}
colnames(amo_yrmt)[2] <- "month"
colnames(amo_yrmt)[3] <- "amo_sample"
amo_yrmt
```

@@

ok

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
Th
```{r pre-install packages, message=FALSE}
packages <- c("packrat", "rworldmap", "dplyr","plyr", "rgdal", "sp", "raster", "tibble", "useful")
source("../src/ipak.R")
ipak(packages)
suppressMessages(ipak)
```

@@

amo_yrmt <- amo_yrmt[order(amo_yrmt$year),]
amo_yrmt
```
rename time to month.. and value to amo_sample
```{r}
colnames(amo_yrmt)[2] <- "month"
colnames(amo_yrmt)[3] <- "amo_sample"
amo_yrmt
```

@@

ok next - AMO value at previous sampling month/year

shift the amo_sample column down by 1...
```{r}
library(useful)
amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
amo_yrmt
```

Note the shifted values are in a new column called amo_sample.Shifted. rename to amo_prev
```{r}
colnames(amo_yrmt)[colnames(amo_yrmt)=="amo_sample.Shifted"] <- "amo_prev"
amo_yrmt
```

```{r}
data_aea <- merge(x = data_aea, y = amoshift[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
data_aea
```

and now - AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)

same process as before - shift row then match

But first!

You need to extract the seasonal values from the amo dataset
```{r}

```

```{r}
amo_yrmt <- shift.column(data = amo_yrmt, columns = "amo_sample", len = 1)
amo_yrmt
```
